---
format: 
  html:
    page-layout: full
editor: visual
---

## Two weeks in Bhutan

In early April 2026 I was very lucky to spend two weeks on holiday in Bhutan. This is a very special country, with a rich culture and beautiful forests for hiking. Below are two maps showing our travel route through Bhutan. The first map gives you an idea of the landscape's topography and on the second map you can zoom in to see each town. Hover over the icons to see which places we visited. We had so many wonderful experiences, but the one that stands out the most was our overnight hike and camping trip. We camped in the hills above Thimphu at an altitude of about 3,700 m, in the snow!

```{r}
#| echo: false
#| message: false
#| output: false

library(tidyverse)
library(sf)
library(tigris) # Access geographic data from the US Census
library(tidygeocoder) # Automated geocoding
library(osrm)
library(ggrepel)
library(scales)
library(ggspatial)
library(leaflet)
library(fontawesome)
library(htmlwidgets)
library(webshot2)
library(htmltools)
```


```{r}
#| echo: false
#| message: false
#| output: false
#| eval: false

stops_address <- tribble(
  ~travel_day, ~place, ~address, ~route_type,
  0, "Paro International Airport", "Paro International Airport, Paro, Bhutan", "driving",
  #1, "Tashi Namgay Resort", "Tashi Namgay Resort, Paro, Bhutan", "driving",
  1, "Rinpung Dzong", "Rinpung, Paro, Bhutan", "walking",
  1, "Tshechu Festival", "Rinpung courtyard, Paro, Bhutan", "driving",
  #2, "Tiger's Nest Road End", "Road end, Paro Taktsang, Paro, Bhutan", "walking",
  2, "Tiger's Nest (3120 m)", "Paro Taktsang, Paro, Bhutan", "walking",
  2, "Taktsang trail", "Road end, Paro Taktsang, Paro, Bhutan", "driving",
  #3, "Tashi Namgay Resort", "Tashi Namgay Resort, Paro, Bhutan", "driving",
  3, "Kila Goenpa Nunnery", "Kila Goenpa, Paro, Bhutan", "walking",
  3, "Chelela Pass (3988 m)", "Chele La Pass, Paro, Bhutan", "driving",
  3, "White Temple", "Lhakhang Karpo, Haa, Bhutan", "driving",
  3, "Sonam Zhidey Resort", "Sonam Zhidey Resort, Haa, Bhutan", "driving",
  4, "ChuZom-Haa Highway1", "ChuZom-Haa Highway, Bhutan","driving",
  4, "ChuZom-Haa Highway2", "ChuZom-Haa Highway, Bhutan","driving",
  #4, "Tango Roadend", "Tango Monastery Roadend, Thimphu, Bhutan", "walking",
  4, "Tango Monastery", "Tango Monastery, Thimphu, Bhutan", "walking",
  4, "Tango Roadend", "Tango Monastery Roadend, Thimphu, Bhutan", "driving",
  4, "Takin Preserve", "Motithang Takin Preserve, Thimphu, Bhutan", "driving",
  #5, "Norkhil Boutique Hotel",     "Norkhil Boutique Hotel, Thimphu, Bhutan", "driving",
  5, "Folk Heritage Museum", "Folk Heritage Museum Kawajangsa, Thimphu, Bhutan", "driving",
  6, "Dochula Pass (3100 m)", "Dochula Pass, Thimphu, Bhutan", "driving",
  6, "Teoprongchu", "Teoprongchu, Bhutan", "walking",
  6, "Chimi Lhakhang (The Temple of Fertility)", "Chimi Lhakhang, Punakha, Bhutan", "walking",
  6, "Teoprongchu", "Teoprongchu, Bhutan", "driving",
  7, "Mo Chhu River, Punakha", "Khamsum Yueli, Namgyal 13001, Bhutan", "walking",
  7, "Khamsum Yulley Namgyal Chorten", "Khamsum Yulley Namgyal Chörten, Thimphu-Punakha Hwy, Bhutan", "walking",
  7, "Mo Chhu River, Punakha", "Khamsum Yueli, Namgyal 13001, Bhutan", "driving",
  7, "Sangchen Dorji Lhuendrup Nunnery", "Sangchhen Dorji Lhuendrup Nunnery, Punakha, Bhutan", "driving",
  #7, "Wangdue Ecolodge", "Wangdue Ecolodge, Damina Village, Ngashigaykha, Rubesa, Wangdue Phodrang, 14001, Bhutan", "driving",
  8, "Wangdue Dzong", "Wangdue Phodrang Dzong, Wangdue Phodrang, Bhutan", "driving",
  8, "Lawala Pass (3250 m)", "Lawala Pass, Dungdungneysa, Bhutan", "driving",
  8, "Gangtey Monastery", "Gangtey Monastery, Phobjikha Valley, Bhutan", "walking",
  8, "Gantye Nature Trail", "Gantye Nature TrailPhobjikha Valley, Bhutan", "walking",
  8, "Phobjikha View Point", "Phobjikha View Point, Phobjikha Valley, Bhutan", "driving",
  8, "Black Necked Crane Visitor Centre", "Black Necked Crane Visitor Centre, Phobjikha, Bhutan", "driving",
  #8, "Gangtey Tent Resort", "Gangtey Tent Resort, Gangtey Phobjikha, Bhutan", "driving",
  9, "Kumbu", "Kumbu, Bhutan", "walking",
  9, "Kaychela Pass (3600 m)", "Kaychela Pass, Bhutan","walking",
  9, "Longtey", "Longtey, Bhutan", "driving",
  #9, "Gangtey Tent Resort", "Gangtey Tent Resort, Gangtey Phobjikha, Bhutan", "driving",
  #10, "Himalayan Keys Forest Resort", "Himalayan Keys Forest Resort, Zhori Zur Lam, Thimphu, Bhutan", "driving",
  10, "TaBar Nye Monastery", "TaBar Nye, Thimphu, Bhutan", "walking",
  10, "Camp (3700 m)", "Thadrana Telecom Tower Junction, Thimphu, Bhutan", "walking",
  10, "Gyalpo Pelzang Peak (4200 m)", "Thadrana Telecom Tower point, Yusipang, Bhutan", "walking",
  11, "Hontsho", "Hontsho Picnic, Hungtsho, Bhutan", "driving")
  #11, "Tashi Namgay Resort", "Tashi Namgay Resort, Paro, Bhutan"


stops_geocoded <- stops_address |>
  geocode(address, method = "osm") |>
  mutate(
    lat = case_when(
      place == "ChuZom-Haa Highway1" ~ 27.31161,
      place == "ChuZom-Haa Highway2" ~ 27.19582,
      place == "Sonam Zhidey Resort" ~ 27.39062,
      place == "Rinpung Dzong" ~ 27.42960,
      #place == "Taktsang trail" ~ 27.47139,
      place == "Taktsang trail" ~27.48180,
      place == "Tango Roadend" ~ 27.50910,
      place == "Tango Monastery" ~ 27.51028,
      #place == "Tshechu Festival" ~ 27.42908,
      place == "Tshechu Festival" ~ 27.425904, 
      place == "Kila Goenpa Nunnery" ~ 27.39064,
      #place == "Khamsum Yulley Namgyal Chorten" ~ 27.62631, My photo GPS seems incorrect
      place == "Khamsum Yulley Namgyal Chorten" ~ 27.63724, 
      #place == "Mo Chhu River, Punakha" ~ 27.62371,
      place == "Mo Chhu River, Punakha" ~ 27.63202, 
      place == "Sangchen Dorji Lhuendrup Nunnery" ~ 27.54797,
      #place == "Wangdue Ecolodge" ~ 27.47045,
      #place == "Lawala Pass" ~ 27.51032,
      place == "Lawala Pass (3250 m)" ~ 27.52489,
      #place == "Gangtey Monastery" ~ 27.47145,
      place == "Gangtey Monastery" ~27.48485, 
      place == "Gantye Nature Trail" ~ 27.48923,
      place == "Phobjikha View Point" ~ 27.47071,
      #place == "Black Necked Crane Visitor Centre" ~ 27.46819,
      #place == "Gangtey Tent Resort" ~ 27.47103,
      place == "Black Necked Crane Visitor Centre" ~27.46456
      #place == "Kumbu" ~ 27.50647,
      place == "Kumbu" ~ 27.51193,
      place == "Kaychela Pass (3600 m)" ~ 27.50639,
      #place == "Longtey" ~ 27.5075,
      place == "Longtey" ~ 27.53060,
      #place == "Himalayan Keys Forest Resort" ~ 27.46988,
      place == "TaBar Nye Monastery" ~ 27.51009,
      place == "Camp (3700 m)" ~ 27.50897,
      place == "Gyalpo Pelzang Peak (4200 m)" ~ 27.50863,
      place == "Hontsho" ~ 27.50726,
      TRUE ~ lat
    ),
    long = case_when(
      place == "ChuZom-Haa Highway1" ~ 89.31698,
      place == "ChuZom-Haa Highway2" ~ 89.49546,
      place == "Sonam Zhidey Resort" ~ 89.40642,
      #place == "Rinpung Dzong" ~ 89.40768,
      #place == "Taktsang trail" ~ 89.36045,
      place == "Taktsang trail" ~ 89.360695,
      place == "Tango Roadend" ~ 89.63318,
      place == "Tango Monastery" ~ 89.62429,
      #place == "Tshechu Festival" ~ 89.40546,
      place == "Tshechu Festival" ~ 89.425298,
      place == "Kila Goenpa Nunnery" ~ 89.36146,
      #place == "Khamsum Yulley Namgyal Chorten" ~ 89.80070,
      place == "Khamsum Yulley Namgyal Chorten" ~ 89.81667,
      #place == "Mo Chhu River, Punakha" ~ 89.80130,
      place == "Mo Chhu River, Punakha" ~ 89.81718,
      place == "Sangchen Dorji Lhuendrup Nunnery" ~ 89.84384,
      place == "Wangdue Ecolodge" ~ 89.89029,
      #place == "Lawala Pass" ~ 90.15212,
      place == "Lawala Pass (3250 m)" ~ 90.17523,
      #place == "Gangtey Monastery" ~ 90.15188,
      place == "Gangtey Monastery" ~ 90.16585,
      place == "Gantye Nature Trail" ~ 90.166495,
      place == "Phobjikha View Point" ~ 90.20001,
      #place == "Black Necked Crane Visitor Centre" ~ 90.15385,
      place == "Black Necked Crane Visitor Centre" ~90.17026
      place == "Gangtey Tent Resort" ~ 90.16570,
      #place == "Kumbu" ~ 90.19599,
      place == "Kumbu" ~ 90.18375,
      place == "Kaychela Pass (3600 m)" ~ 90.19788,
      #place == "Longtey" ~ 90.1964,
      place == "Longtey" ~ 90.2152,
      place == "Himalayan Keys Forest Resort" ~ 89.62807,
      place == "TaBar Nye Monastery" ~ 89.67089,
      place == "Camp (3700 m)" ~ 89.71630,
      place == "Gyalpo Pelzang Peak (4200 m)" ~ 89.71298,
      place == "Hontsho" ~ 89.71271,
      TRUE ~ long
    )
  )
```

```{r}
#| echo: false
#| message: false
#| output: false

stops_geocoded <- read_csv("data/stops_geocoded.csv")

all_stops_unique <- stops_geocoded |>
  distinct()

stops_sf <- stops_geocoded |>
  st_as_sf(coords = c("long", "lat"), crs = st_crs("EPSG:4326"))

towns <- tribble(
  ~travel_day, ~town,        ~hotel,                      ~address,
  1,    "Paro",       "Tashi Namgay Resort",        "Tashi Namgay Resort, Paro, Bhutan",
  2,    "Paro",       "Tashi Namgay Resort",        "Tashi Namgay Resort, Paro, Bhutan",
  3,    "Haa",        "Sonam Zhidey Resort",        "Sonam Zhidey Resort, Haa, Bhutan",
  4,    "Thimphu",    "Norkhil Boutique Hotel",     "Norkhil Boutique Hotel, Thimphu, Bhutan",
  5,    "Thimphu",    "Norkhil Boutique Hotel",     "Norkhil Boutique Hotel, Thimphu, Bhutan",
  6,    "Punakha",    "Dumra Farm Resort",          "Dumra Farm Resort, Punakha, Bhutan",
  7,    "Wangdue",    "Wangdue Ecolodge",           "Wangdue Ecolodge, Damina Village, Ngashigaykha, Rubesa, Wangdue Phodrang, 14001, Bhutan",
  8,    "Gangtey",    "Gangtey Tent Resort",        "Gangtey Tent Resort, Gangtey Phobjikha, Bhutan",
  9,    "Gangtey",    "Gangtey Tent Resort",        "Gangtey Tent Resort, Gangtey Phobjikha, Bhutan",
  #10, "Thimphu", "Himalayan Keys Forest Resort", "Himalayan Keys Forest Resort, Zhori Zur Lam, Thimphu, Bhutan"
)

towns_geocoded <- towns |>
  #geocode(address, method = "osm") |>
  mutate(
    lat = case_when(hotel == "Tashi Namgay Resort" ~ 27.39000,
                    hotel == "Sonam Zhidey Resort" ~ 27.39062,
                    hotel == "Norkhil Boutique Hotel" ~ 27.46915,
                    hotel == "Dumra Farm Resort" ~ 27.58620,
                    hotel == "Wangdue Ecolodge" ~ 27.47045,
                    #hotel == "Gangtey Tent Resort" ~ 27.47103,
                    hotel == "Gangtey Tent Resort" ~ 27.48603),
                    #hotel == "Himalayan Keys Forest Resort" ~ 27.46988),
    long = case_when(hotel == "Tashi Namgay Resort" ~ 89.40642,
                     hotel == "Sonam Zhidey Resort" ~ 89.27509,
                     hotel == "Norkhil Boutique Hotel" ~ 89.62667,
                     hotel == "Dumra Farm Resort" ~ 89.84649,
                     hotel == "Wangdue Ecolodge" ~ 89.89029,
                     hotel == "Gangtey Tent Resort" ~ 90.15620))
                     #hotel == "Himalayan Keys Forest Resort" ~ 89.62807)
    
                    
all_towns_unique <- towns_geocoded |>
  distinct()

all_stops_cat <- all_stops_unique |>
  mutate(category = case_when(
    grepl("Airport", place, ignore.case = TRUE) ~ "airport",
    #grepl("Nest|Hike|Trek|Chelela", place, ignore.case = TRUE) ~ "hiking",
    grepl("Camp", place, ignore.case = TRUE) ~ "camping",
    grepl("River", place, ignore.case = TRUE) ~ "rafting",
    grepl("Dzong|Temple|Monastery|Nunnery|Goenpa|Nest|Chorten", place, ignore.case = TRUE) ~ "temple",
    grepl("View|Black|Preserve", place, ignore.case = TRUE) ~ "wildlife",
    #grepl("Resort|Ecolodge", place, ignore.case = TRUE) ~ "hotel",
    grepl("Heritage|Festival", place, ignore.case = TRUE) ~ "culture",
    grepl("Pass", place, ignore.case = TRUE) ~ "pass",
    TRUE ~ "other"
  ))
```

```{r}
#| echo: false
#| message: false
#| output: false

# Function that routes based on type#

get_route <- function(src_coords, dst_coords, route_type) {
  if (route_type == "walking") {
    line <- sf::st_sfc(
      sf::st_linestring(rbind(src_coords, dst_coords)),
      crs = 4326
    )
    dist_km <- as.numeric(sf::st_length(line)) / 1000
    
    sf::st_sf(
      src      = paste(src_coords, collapse = ","),
      dst      = paste(dst_coords, collapse = ","),
      duration = NA_real_,
      distance = dist_km,
      geometry = line
    )
  } else {
    osrmRoute(
      src          = src_coords,
      dst          = dst_coords,
      osrm.profile = "car"
    )
  }
}

```

```{r}
#| echo: false
#| message: false
#| output: false

routes_geocoded <- stops_sf |>
  mutate(
    origin_place      = place,
    destination_place = lead(place),
    src_long          = st_coordinates(geometry)[, 1],
    src_lat           = st_coordinates(geometry)[, 2],
    dest_long         = lead(st_coordinates(geometry)[, 1]),
    dest_lat          = lead(st_coordinates(geometry)[, 2])
  ) |>
  filter(!is.na(destination_place)) |>
  rowwise() |>
  mutate(
    route = list(get_route(
      src_coords = c(src_long, src_lat),
      dst_coords = c(dest_long, dest_lat),
      route_type = route_type
    ))
  )

routes_sf <- routes_geocoded |>
  ungroup() |>
  sf::st_drop_geometry() |>                 # drop POINT geometry from stops_sf
  dplyr::select(-route) |>                  # keep the attributes (day, route_type, etc.)
  dplyr::bind_cols(
    do.call(rbind, routes_geocoded$route) |>  # stack all the per-segment sf frames
      sf::st_as_sf() |>
      dplyr::select(geometry)                  # keep just the LINESTRING column
  ) |>
  sf::st_as_sf()

```

```{r}
#| echo: false
#| message: false
#| output: false

fa_svg_icon <- function(name, fill = "black", size = 16) {
  url <- sprintf(
    "https://cdn.jsdelivr.net/npm/@fortawesome/fontawesome-free@6.5.1/svgs/solid/%s.svg",
    name
  )
  svg <- paste(readLines(url, warn = FALSE), collapse = "")
  # Inject the fill into the <svg> tag
  svg_colored <- sub("<svg ", sprintf('<svg fill="%s" ', fill), svg)
  data_uri <- paste0(
    "data:image/svg+xml;utf8,",
    utils::URLencode(svg_colored, reserved = TRUE)
  )
  makeIcon(iconUrl = data_uri, iconWidth = size, iconHeight = size)
}


icons_svg <- iconList(
  airport = fa_svg_icon("plane", "white"),
  camping  = fa_svg_icon("tent",   "white"),
  culture  = fa_svg_icon("camera", "white"),
  rafting  = fa_svg_icon("water",  "white"),
  pass = fa_svg_icon("mountain", "white"),
  temple   = fa_svg_icon("vihara", "white"),
  wildlife = fa_svg_icon("crow",   "white")
  )

icons_svg <- iconList(
  airport = fa_svg_icon("plane", "lightgrey"),
  camping  = fa_svg_icon("tent",   "lightgrey"),
  culture  = fa_svg_icon("camera", "lightgrey"),
  rafting  = fa_svg_icon("water",  "lightgrey"),
  pass = fa_svg_icon("mountain", "lightgrey"),
  temple   = fa_svg_icon("vihara", "lightgrey"),
  wildlife = fa_svg_icon("crow",   "lightgrey")
  )

route_legend <- tags$div(
  style = paste(
    "background: rgba(250, 250, 250);",
    "color: #1E2019;", 
    "border: 2px solid #777;",
    "border-top: none;",          
    "padding: 2px 8px 4px;",
    "font-size: 11px;",
    "line-height: 22px;"
  ),
  tags$b("Route type"), tags$br(),
  tags$span(
    style = "display:inline-block; width:36px; height:0; border-top:4px solid #AB6AD5; vertical-align:middle; margin-right:6px;"
  ), "Driving", tags$br(),
  tags$span(
    style = "display:inline-block; width:36px; height:0; border-top:3px dashed #DC5FBD; vertical-align:middle; margin-right:6px;"
  ), "Walking"
)
```

```{r}
#| echo: false
#| message: false
#| output: false

bhutan_map1 <- leaflet() |>
  #addProviderTiles("Stadia.StamenTerrainBackground")  |> 
  addProviderTiles("Esri.WorldImagery")  |>
  #addProviderTiles("OpenStreetMap.Mapnik")  |>
  addPolylines(
    data      = routes_sf |> filter(route_type == "driving"),
    #color     = "#C6C1F0FF",
     color     = "#AB6AD5FF",
    weight    = 4,
    opacity   = 0.8,
    dashArray = NULL
      ) |>
  addPolylines(
    data      = routes_sf |> filter(route_type == "walking"),
    #color     = "#F498B6FF",
    color     = "#DC5FBDFF",
    weight    = 3,
    opacity   = 0.8,
    dashArray = "1, 5"
  ) |> 
  addMarkers(
    data  = all_stops_cat |> filter(category %in% names(icons_svg)),
    lat   = ~lat, lng = ~long,
    icon  = ~icons_svg[category],
    label = ~paste0("Day ", as.integer(travel_day), ": ", place)
    #popup = ~paste0("<b>", place, "Day: ", as.integer(travel_day))
  ) |> 
  addLabelOnlyMarkers(
    data = all_towns_unique,
    lat = ~lat,
    lng = ~long,
    label = ~town,
    labelOptions = labelOptions(
      noHide = TRUE,
      #direction = "top",
      textOnly = TRUE,
      style = list(
        "font-weight" = "bold",
        "font-size"   = "14px",
        "color"       = "lightgrey",
        "text-shadow" = "1px 1px 2px black"
      )
    )) |> 
  addScaleBar(position = "bottomleft", options = scaleBarOptions(maxWidth = 100, 
                                                                 updateWhenIdle = TRUE,
                                                                 imperial = FALSE)) |> 
  addControl(html = route_legend, position = "bottomright", className = "route-legend")|> 
  htmlwidgets::prependContent(
  tags$style(HTML(".route-legend { padding: 0 !important; }"))
)

```

```{r}
#| echo: false
#| message: false

bhutan_map1

```

<br>

```{r}
#| echo: false
#| message: false
#| output: false

icons_svg <- iconList(
  airport = fa_svg_icon("plane", "black"),
  camping  = fa_svg_icon("tent",   "black"),
  culture  = fa_svg_icon("camera", "black"),
  rafting  = fa_svg_icon("water",  "black"),
  pass = fa_svg_icon("mountain", "black"),
  temple   = fa_svg_icon("vihara", "black"),
  wildlife = fa_svg_icon("crow",   "black")
  )

route_legend <- tags$div(
  style = paste(
    "background: rgba(250, 250, 250);",
    "color: #1E2019;", 
    "border: 2px solid #777;",
    "border-top: none;",          
    "padding: 2px 8px 4px;",
    "font-size: 11px;",
    "line-height: 22px;"
  ),
  tags$b("Route type"), tags$br(),
  tags$span(
    style = "display:inline-block; width:36px; height:0; border-top:4px solid #C6C1F0; vertical-align:middle; margin-right:6px;"
  ), "Driving", tags$br(),
  tags$span(
    style = "display:inline-block; width:36px; height:0; border-top:3px dashed #F498B6; vertical-align:middle; margin-right:6px;"
  ), "Walking"
)

bhutan_map2 <- leaflet() |>
  #addProviderTiles("Stadia.StamenTerrainBackground")  |> 
  #addProviderTiles("Esri.WorldImagery")  |>
  addProviderTiles("OpenStreetMap.Mapnik")  |>
  addPolylines(
    data      = routes_sf |> filter(route_type == "driving"),
    #color     = "#C6C1F0FF",
     color     = "#AB6AD5FF",
    weight    = 4,
    opacity   = 0.8,
    dashArray = NULL
      ) |>
  addPolylines(
    data      = routes_sf |> filter(route_type == "walking"),
    #color     = "#F498B6FF",
    color     = "#DC5FBDFF",
    weight    = 3,
    opacity   = 0.8,
    dashArray = "1, 5"
  ) |> 
  addMarkers(
    data  = all_stops_cat |> filter(category %in% names(icons_svg)),
    lat   = ~lat, lng = ~long,
    icon  = ~icons_svg[category],
    label = ~paste0("Day ", as.integer(travel_day), ": ", place)
    #popup = ~paste0("<b>", place, "Day: ", as.integer(travel_day))
  ) |> 
  addLabelOnlyMarkers(
    data = all_towns_unique,
    lat = ~lat,
    lng = ~long,
    label = ~town,
    labelOptions = labelOptions(
      noHide = TRUE,
      #direction = "top",
      textOnly = TRUE,
      style = list(
        "font-weight" = "bold",
        "font-size"   = "14px",
        "color"       = "lightgrey",
        "text-shadow" = "1px 1px 2px black"
      )
    )) |> 
  addScaleBar(position = "bottomleft", options = scaleBarOptions(maxWidth = 100, 
                                                                 updateWhenIdle = TRUE,
                                                                 imperial = FALSE)) |> 
  addControl(html = route_legend, position = "bottomright", className = "route-legend")|> 
  htmlwidgets::prependContent(
  tags$style(HTML(".route-legend { padding: 0 !important; }"))
)

```

```{r}
#| echo: false
#| message: false

bhutan_map2
```

<br>

Note: I used straight lines for the walks as the Open Source Routing Machine  did not work along the walking tracks. 
